### 2025-08-29 manuscript/manuscript.qmd
Added more to keyword analysis

abstract: "Academia suffers from issues regarding equity, with one  crucial aspect being academic authorship. Many studies have shown that there is a gender gap in publishing, where women are often underrepresented. As there are differences across disciplines, the goal of this study is to provide a current overview the gender distribution of authors within music psychology. A total of 3,373 papers published between 2000 and 2025 across five journals were analyzed with respect to author gender and authorship role (first, last, solo, coauthor). In addition, patterns related to citations, open-access publishing, and keywords were explored across gender. Overall, female authors were in the minority, accounting for 40.2% of all authors. However, the distribution of authorship roles diverged notably from this overall proportion: women were more likely to be first authors (OR = 1.41), equally likely to be solo authors (odds ratio, OR = 1.11) or coauthors (OR = 1.00), but less likely to be last authors (OR = 0.73). These patterns have become more pronounced over the past 15 years. No substantial gender differences in citation counts were observed. Analyses of geographical variation revealed clear cross-country differences, potentially reflecting both the status of women and the historical development of the discipline in particular regions. Finally, an examination of keywords suggested broad thematic preferences differing by gender. While overall some of this work shows that the field of music psychology is more balanced than some other field, especially in more STEM disciplines, providing such a picture hopefully inspires discussions and initiatives to promote fairness and more equity in academia."  
keywords          : "gender, transparency, music psychology, open science, meta-research"
wordcount         : "5555"
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no
mask              : no
csl: apa7.csl
class             : "man"
bibliography: references.bib
format:
  html: default
---

>>>>

abstract: "Academia suffers from issues regarding equity, where academic authorship is one aspect where inequality occurs. Many studies have shown that there is a gender gap in publishing, where women are often underrepresented. As there are differences across disciplines, the goal of this study is to provide a current overview the gender distribution of authors within music psychology. A total of 3,373 papers published between 2000 and 2025 across five journals were analyzed with respect to author gender and authorship role (first, last, solo, coauthor). In addition, patterns related to citations, open-access publishing, and keywords were explored across gender. Overall, female authors were in the minority, accounting for 40.2% of all authors. However, the distribution of authorship roles diverged notably from this overall proportion: women were more likely to be first authors (OR = 1.41), equally likely to be solo authors (odds ratio, OR = 1.11) or coauthors (OR = 1.00), but less likely to be last authors (OR = 0.73). These patterns have become more pronounced over the past 15 years. No substantial gender differences in citation counts were observed. Analyses of geographical variation revealed clear cross-country differences, potentially reflecting both the status of women and the historical development of the discipline in particular regions. Finally, an examination of keywords suggested broad thematic preferences differing by gender. While overall some of this work shows that the field of music psychology is more balanced than some other fields, especially in more STEM disciplines, providing such a picture hopefully inspires discussions and initiatives to promote fairness and more equity in academia."

keywords          : "gender, transparency, music psychology, open science, meta-research"
wordcount         : "5555"
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no
mask              : no
csl: apa7.csl
class             : "man"
bibliography: references.bib
format:
  html: default
---
